---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Sam Leuthold"
date: "2-7-2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)
library(pdftools)
library(ggpubr)
library(ggthemes)
```

# Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r, message = F}

site_url <- 'https://snowstudies.org/archived-data/'

webpage <- read_html(site_url)

links <- webpage %>%
           html_nodes('a') %>%
           .[grepl('forcing',.)] %>%
           html_attr('href')

```

2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r, results = F}

labels <- links %>% 
            str_split_fixed("/", n = 8)

dataset <- labels[,8] 

file.names <- paste0('data/',dataset)

map2(links[1:2],
     file.names[1:2],
     download.file)

```

3. Write a custom function to read in the data and append a site column to the data. 

```{r, message = F}

headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
            readr::read_lines(.) %>%
            trimws(.) %>%
            str_split_fixed(.,'\\.',2) %>%
            .[,2] %>%
            .[1:26] %>%
            str_trim(side = "left")

## I'd be interested to see the better Tidy version of this function. I messed with it for a while, but this ended up being the easiest way for me. 

Meteorological_Data_Download <- function(site.name){
                                  
                                  file.index <- which(grepl(site.name, file.names) == T)
                                  
                                  file.path <- file.names[file.index]
                                    
                                  output <- read_fwf(file.path, 
                                                     col_positions = fwf_empty(file.path), 
                                                     show_col_types = F)
                                  
                                  colnames(output) <- headers
    
                                  output <- output[,1:14]
                                  
                                  output <- output %>%
                                              mutate(Site = site.name, .before = year)
                                  
                                  assign(paste0(site.name, ".data2"), output, envir = parent.frame())
}

```

4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r, message = F}

sites <- c("SASP", "SBSP")

climate.data <- map_dfr(.x = sites[1:2], 
                        .f = Meteorological_Data_Download)


knitr::kable(summary(climate.data))

```


5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

```{r, message = F}

mean.temp <- climate.data %>% 
              filter(year != 2003) %>%
              group_by(Site, year) %>%
              summarise(Mean_Temp = mean(`air temp [K]`, na.rm = TRUE))


ggplot(data = mean.temp, aes(x = year, y = Mean_Temp)) +
  geom_line(aes(color = Site), 
            size = 1.5) +
  geom_point(shape = 21,
             size = 4,
             aes(fill = Site)) +
  scale_fill_colorblind() +
  scale_color_colorblind() +
  scale_x_continuous(name = "Year", 
                     breaks = seq(2003, 2011)) +
  ylab("Average Annual Air Temperature (K)") +
  theme_classic2() +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_text(color = "black"),
        legend.position = c(0.85,0.15))
  


```

6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html

```{r, message = FALSE}


Avg_Temp_Plotter <- function(Year){
  
  temp.temp.data <- climate.data %>%
    group_by(Site, year, month) %>%
    summarise(Monthly_Mean = mean(`air temp [K]`)) %>%
    filter(year == Year)
  
  
  temp.plot <- ggplot(data = temp.temp.data, aes(x = month, y = Monthly_Mean)) +
  geom_line(aes(color = Site), 
            size = 1.5) +
  geom_point(shape = 21,
             size = 4,
             aes(fill = Site)) +
  scale_fill_colorblind() +
  scale_color_colorblind() +
  scale_x_continuous(name = "Year", 
                     breaks = seq(1,12)) +
  scale_y_continuous(name = "Average Monthly Air Temperature (K)",
                    limits = c(240, 300)) +
    ggtitle(label = Year) +
  theme_classic2() +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_text(color = "black"),
        legend.position = c(0.85,0.15))
  
  print(temp.plot)
  
}

years <- seq(2005, 2010)

for (i in 1:length(years)){
  
  Avg_Temp_Plotter(years[i])
  
}

```

Bonus: Make a plot of average daily precipitation by day of year (averaged across all available years). Color each site. 

``` {r, message = FALSE}

daily.precipitation <-  climate.data %>% 
                          mutate(Date = paste0(day,"/", month, "/", year)) %>%
                          mutate(Date = as.Date(Date, format = "%d/%m/%Y")) %>%
                          mutate(DOY = yday(Date)) %>%
                          mutate(Precip_mm = `precip [kg m-2 s-1]` * 86400) %>%
                          group_by(Site, DOY) %>%
                          summarize(Mean_Precip = mean(Precip_mm))

ggplot(daily.precipitation[daily.precipitation$Site == "SASP",], aes(x = DOY, y = Mean_Precip)) +
  geom_col(color = "black",
           fill = "lightblue") +
  scale_x_continuous(name = "Julian Day", 
                     limits = c(-2, 367),
                     breaks = seq(0, 360, 30),
                     expand = c(0,0)) +
   scale_y_continuous(name = "Mean Daily Precip (mm)",
                     limits = c(0, 15),
                     expand = c(0,0)) + 
  theme_classic2() +
  theme(panel.border = element_rect(fill = NA),
  axis.text = element_text(color = "black"),
  legend.position = c(0.85,0.15))

```

Bonus #2: Use a function and for loop to create yearly plots of precipitation by day of year. Color each site. 

```{r, message = FALSE, warning=FALSE}

Yearly_Precip <- function(input.year){
              
                  precip.out <- climate.data %>%
                                  filter(year == input.year) %>% 
                                  filter(Site == "SASP") %>% 
                                  mutate(Date = paste0(day,"/", month, "/", year)) %>%
                                  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) %>%
                                  mutate(DOY = yday(Date)) %>%
                                  mutate(Precip_mm = `precip [kg m-2 s-1]` * 86400) %>%
                                  group_by(DOY) %>%
                                  summarize(Mean_Precip = mean(Precip_mm))

}


for(i in 1:length(unique(climate.data$year))){
  
  annual.precip <- Yearly_Precip(unique(climate.data$year)[i])

  plot <- ggplot(data = annual.precip, aes(x = DOY, y = Mean_Precip)) +
           geom_col(color = "black",
                    fill = "lightblue") +
           scale_x_continuous(name = "Julian Day", 
                              limits = c(-2, 367),
                              breaks = seq(0, 360, 30),
                              expand = c(0,0)) +
          scale_y_continuous(name = "Mean Daily Precip (mm)",
                             limits = c(0, 50),
                             expand = c(0,0)) + 
           ggtitle(label = unique(climate.data$year)[i]) +
           theme_classic2() +
           theme(panel.border = element_rect(fill = NA),
                 axis.text = element_text(color = "black"),
                 legend.position = c(0.85,0.15))
  
  print(plot)
  
}
  
```
